---
title: "3. Fine-Gray Analysis"
author: "SeuljiMoon"
date: "2023-08-28"
output: html_document
---

### install.packages
```{r global-options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
if(!require("ggplot2")) install.packages("ggplot2")
if(!require("tableone")) install.packages("tableone")
if(!require("data.table")) install.packages("data.table")
if(!require("ggpubr")) install.packages("ggpubr")
if(!require("gridExtra")) install.packages("gridExtra")
if(!require("cobalt")) install.packages("cobalt")
if(!require("MatchIt")) install.packages("MatchIt")
if(!require("magrittr")) install.packages("magrittr")
if(!require("moonBook")) install.packages("moonBook")
if(!require("dplyr")) install.packages("dplyr")
if(!require("parallel")) install.packages("parallel")
if(!require("survival")) install.packages("survival")
if(!require("cmprsk")) install.packages("cmprsk")
if(!require("lubridate")) install.packages("lubridate")
if(!require("survplot")) install.packages("survplot")
     
library(data.table); library(dplyr);
library(magrittr);library(parallel);library(ggplot2);library(tableone)
library(survival);library(survminer);library(cmprsk)
library(lubridate)
library(survplot)
library(moonBook)
```

### Data manipulation
```{r}
source("2. Cumulative_incidence_plot.Rmd")
```


### Estimate cumulative incidence and it's 95% CI at time points 
 - Estimate cumulative incidence functions from competing risks data and test equality across groups
 - Using `cuminc` function in **cmprsk** package
 - Status value : 0 - censoring / 1 - event / 2 - competing risk
 - Group : 1 - Case (Multiple myeloma) / 2 - Control

```{r}
ci.model0<-cuminc(ftime=frac_entire$surv_years,
                 fstatus=frac_entire$event3,
                 group=frac_entire$c90_yn,cencode = 0)

# frac event
citable_control0 <- ci.model0$`0 1` %>% as.data.table %>% .[,1:3]
citable_control0$std <- sqrt(citable_control0$var)
citable_case0<-ci.model0$`1 1` %>% as.data.table %>% .[,1:3]
citable_case0$std <- sqrt(citable_case0$var)
  
# competing evnet
citable_control_d0<-ci.model0$`0 2` %>% as.data.table %>% .[,1:3]
citable_control_d0$std <- sqrt(citable_control_d0$var)
citable_case_d0<-ci.model0$`1 2` %>% as.data.table %>% .[,1:3]
citable_case_d0$std <- sqrt(citable_case_d0$var)
  
tt <- c()
for (i in 1:8) {
  th <- data.table(
    year = i,
    case = paste0(round(as.numeric(citable_case0[time>i][1,2]),3), "(", 
                  (round(as.numeric(citable_case0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_case0[time>i][1,4]),3)), "-",
                  (round(as.numeric(citable_case0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_case0[time>i][1,4]),3)), ")") ,
    control = paste0(round(as.numeric(citable_control0[time>i][1,2]),3), "(", 
                     (round(as.numeric(citable_control0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_control0[time>i][1,4]),3)), "-",
                     (round(as.numeric(citable_control0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_control0[time>i][1,4]),3)), ")") ,
    p_val = ifelse(ci.model0$Test[1,2]<0.001,"<0.001",round(ci.model0$Test[1,2],3)),
    
    case_death = paste0(round(as.numeric(citable_case_d0[time>i][1,2]),3), "(", 
                        (round(as.numeric(citable_case_d0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_case_d0[time>i][1,4]),3)), "-",
                        (round(as.numeric(citable_case_d0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_case_d0[time>i][1,4]),3)), ")") ,
    control_death = paste0(round(as.numeric(citable_control_d0[time>i][1,2]),3), "(", 
                     (round(as.numeric(citable_control_d0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_control_d0[time>i][1,4]),3)), "-",
                     (round(as.numeric(citable_control_d0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_control_d0[time>i][1,4]),3)), ")") ,
    p_val = ifelse(ci.model0$Test[1,2]<0.001,"<0.001",round(ci.model0$Test[1,2],3))
    
  )

tt <- rbind(tt,th)
}

```


Apply 1-year landmark 
```{r}
frac_1year <- frac_entire[frac_entire$surv_years >= 1,]
frac_1year <- as.data.table(frac_1year)
```

```{r}
ci.model0<-cuminc(ftime=frac_1year$surv_years,
                  fstatus=frac_1year$event3,
                  group=frac_1year$c90_yn,cencode = 0)


# frac event
citable_control0 <- ci.model0$`0 1` %>% as.data.table %>% .[,1:3]
citable_control0$std <- sqrt(citable_control0$var)
citable_case0<-ci.model0$`1 1` %>% as.data.table %>% .[,1:3]
citable_case0$std <- sqrt(citable_case0$var)

# competing evnet
citable_control_d0<-ci.model0$`0 2` %>% as.data.table %>% .[,1:3]
citable_control_d0$std <- sqrt(citable_control_d0$var)
citable_case_d0<-ci.model0$`1 2` %>% as.data.table %>% .[,1:3]
citable_case_d0$std <- sqrt(citable_case_d0$var)

tt <- c()
for (i in 1:8) {
  th <- data.table(
    year = i,
    case = paste0(round(as.numeric(citable_case0[time>i][1,2]),3), "(", 
                  (round(as.numeric(citable_case0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_case0[time>i][1,4]),3)), "-",
                  (round(as.numeric(citable_case0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_case0[time>i][1,4]),3)), ")") ,
    control = paste0(round(as.numeric(citable_control0[time>i][1,2]),3), "(", 
                     (round(as.numeric(citable_control0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_control0[time>i][1,4]),3)), "-",
                     (round(as.numeric(citable_control0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_control0[time>i][1,4]),3)), ")") ,
    p_val = ifelse(ci.model0$Test[1,2]<0.001,"<0.001",round(ci.model0$Test[1,2],3)),
    
    case_death = paste0(round(as.numeric(citable_case_d0[time>i][1,2]),3), "(", 
                        (round(as.numeric(citable_case_d0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_case_d0[time>i][1,4]),3)), "-",
                        (round(as.numeric(citable_case_d0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_case_d0[time>i][1,4]),3)), ")") ,
    control_death = paste0(round(as.numeric(citable_control_d0[time>i][1,2]),3), "(", 
                           (round(as.numeric(citable_control_d0[time>i][1,2]),3) - 1.96*round(as.numeric(citable_control_d0[time>i][1,4]),3)), "-",
                           (round(as.numeric(citable_control_d0[time>i][1,2]),3) + 1.96*round(as.numeric(citable_control_d0[time>i][1,4]),3)), ")") ,
    p_val = ifelse(ci.model0$Test[1,2]<0.001,"<0.001",round(ci.model0$Test[1,2],3))
    
  )
  
  tt <- rbind(tt,th)
  
}
tt
```



### Estimate cumulative incidence and it's 95% CI at time points - Fracture subtypes 
```{r}
tt <- c()
for (i in c("spine","hip","upper")) {
  for (j in c(0,1)) {
    
    df <- frac_entire[frac_entire$surv_years >= j ,] # landmark analysis
    df <- as.data.table(df)
    
    ci.model0<-with(df,cuminc(ftime=surv_years,
                      fstatus=get(paste0("event3_",i)),
                      group=c90_yn,cencode = 0))
    
    # frac event
    citable_control0 <- ci.model0$`0 1` %>% as.data.table %>% .[,1:3]
    citable_control0$std <- sqrt(citable_control0$var)
    citable_case0<-ci.model0$`1 1` %>% as.data.table %>% .[,1:3]
    citable_case0$std <- sqrt(citable_case0$var)
    
    # competing evnet
    citable_control_d0<-ci.model0$`0 2` %>% as.data.table %>% .[,1:3]
    citable_control_d0$std <- sqrt(citable_control_d0$var)
    citable_case_d0<-ci.model0$`1 2` %>% as.data.table %>% .[,1:3]
    citable_case_d0$std <- sqrt(citable_case_d0$var)
    
    
    for (ii in 1:8) {
      th <- data.table(
        dfset = paste0(i," subtype & landmark ",j,"year"),
        year = ii,
        
        case = paste0(round(as.numeric(citable_case0[time>ii][1,2]),3), "(", 
                      (round(as.numeric(citable_case0[time>ii][1,2]),3) - 1.96*round(as.numeric(citable_case0[time>ii][1,4]),3)), "-",
                      (round(as.numeric(citable_case0[time>ii][1,2]),3) + 1.96*round(as.numeric(citable_case0[time>ii][1,4]),3)), ")") ,
        control = paste0(round(as.numeric(citable_control0[time>ii][1,2]),3), "(", 
                         (round(as.numeric(citable_control0[time>ii][1,2]),3) - 1.96*round(as.numeric(citable_control0[time>ii][1,4]),3)), "-",
                         (round(as.numeric(citable_control0[time>ii][1,2]),3) + 1.96*round(as.numeric(citable_control0[time>ii][1,4]),3)), ")") ,
        p_val = ifelse(ci.model0$Test[1,2]<0.001,"<0.001",round(ci.model0$Test[1,2],3)),
        
        case_death = paste0(round(as.numeric(citable_case_d0[time>ii][1,2]),3), "(", 
                            (round(as.numeric(citable_case_d0[time>ii][1,2]),3) - 1.96*round(as.numeric(citable_case_d0[time>ii][1,4]),3)), "-",
                            (round(as.numeric(citable_case_d0[time>ii][1,2]),3) + 1.96*round(as.numeric(citable_case_d0[time>ii][1,4]),3)), ")") ,
        control_death = paste0(round(as.numeric(citable_control_d0[time>ii][1,2]),3), "(", 
                               (round(as.numeric(citable_control_d0[time>ii][1,2]),3) - 1.96*round(as.numeric(citable_control_d0[time>ii][1,4]),3)), "-",
                               (round(as.numeric(citable_control_d0[time>ii][1,2]),3) + 1.96*round(as.numeric(citable_control_d0[time>ii][1,4]),3)), ")") ,
        p_val = ifelse(ci.model0$Test[1,2]<0.001,"<0.001",round(ci.model0$Test[1,2],3))
        
      )
      
      tt <- rbind(tt,th)
    }
  }
}

tt


write.csv(tt, "~\\fracture_subgroup_cuminc.csv")
```

